---
title: "Simulation Methods"
author: "Domenic Di Francesco (CEng)"
date: "July 2018"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
  powerpoint_presentation: default
subtitle: For Structural Integrity Assessment
transition: slowest
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggforce); library(tidyverse); library(extrafont); loadfonts(device='win'); library(lhs); library(fitdistrplus); library(DomDF)
```

## Contents

  - Requirement for Simulation Methods
  
  - Structural Reliability Example

  - Monte Carlo Simulation

  - Latin Hypercube Sampling

  - Importance Sampling

<div class="notes">
This is my *note*.

- It can contain markdown
- like this list

For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

</div>

## Why Simulation Methods?

- Variability of Parameters

```{r, echo=FALSE, fig.width=4, fig.height=4, warning=FALSE}
LNs <- function(m,s) sqrt(log(1 + s^2 / m^2)); LNm <- function(m,s) log(m) - 0.5 * LNs(m,s)^2
Gscale <- function(s) (1 / pi) * s * sqrt(6); Glocation <- function(m,s) m - gamma * Gscale(s)

set.seed(1008); binWidth <- 25
muS <- 300; sigmaS <- 50; N <- 30; muSLN <- LNm(muS, sigmaS); sigmaSLN <- LNs(muS, sigmaS)
S <- rlnorm(n = N, meanlog = muSLN, sdlog = sigmaSLN)

minX <- 0; maxX <- 500; binWidth <- 25

dF <- data.frame(S)

p <- 0.05; lB <- qlnorm(p, meanlog = muSLN, sdlog = sigmaSLN)
model <- fitdist(data = S, distr = 'lnorm')#; summary(model)
test <- ks.test(S, 'plnorm')#; test$p.value

ggplot(dF)+
  geom_histogram(aes(x = S), col='grey25', fill='grey75', binwidth = binWidth)+
  geom_vline(xintercept = lB, linetype=2)+
  geom_label(aes(x=200, y=3, label='5% Lower Bound, \n225.4 MPa'), family='Trebuchet MS', 
             size=3)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  scale_x_continuous(limits = c(minX, maxX), name = 'Yield Strength, MPa')+
  scale_y_continuous(name = 'Count')+
  theme(text = element_text(size = 9), 
        axis.text.y = element_blank())

ggplot(dF)+
  geom_histogram(aes(x = S, y=..density..), col='grey25', fill='grey75', binwidth = binWidth)+
  stat_function(geom='line', fun = dlnorm, args = list(meanlog = muSLN, sdlog = sigmaSLN), linetype=2)+
  geom_label(aes(x=125, y=0.006, label='Lognormal \nApproximation \n ~ LN(300, 50)'), family='Trebuchet MS', 
             size=3)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  scale_x_continuous(limits=c(minX, maxX), name='Yield Strength, MPa')+
  scale_y_continuous(name = 'Probability Density')+
  theme(text = element_text(size = 9), 
        axis.text.y = element_blank())


```

## Structural Reliability Example

\[
 \small{\Pr(Fail) = \int_{g(u) \le 0} g(u) \cdot du}
\]

```{r echo = FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, warning=FALSE}

mu.R <- 180; sigma.R <- 15; mu.S <- 100; sigma.S <- 20
mu.M <- mu.R - mu.S; sigma.M <- sqrt(sigma.R^2 + sigma.S^2)

ggplot(NULL)+
  stat_function(geom = 'density', fun = dnorm, args = list(mean = mu.R, sd = sigma.R), fill='grey', aes(alpha = ' Resistance  '))+
  stat_function(geom = 'density', fun = dnorm, args = list(mean = mu.S, sd = sigma.S), fill='grey', aes(alpha = ' Load  '))+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  scale_x_continuous(limits = c(-50, 300), name = 'Stress or Strength, MPa')+
  scale_y_continuous(name = 'Probability Density')+
  theme(legend.title = element_blank())+theme(legend.position = 'top')+
  theme(text=element_text(size=9))
```

## Structural Reliability Example

\[
  \small{M \sim N \Big(\mu=\mu_R-\mu_s, \quad \sigma = \sqrt{ \sigma_R^2 + \sigma_S^2} \: \Big) }
\]

\[
 \small{\Pr(Fail) = \Pr(M<0)}
\]

```{r echo = FALSE, fig.align='center', fig.width=4, fig.height=4, warning=FALSE}
ggplot(NULL)+
  stat_function(geom = 'density', fun=dnorm, args = list(mean = mu.M, sd = sigma.M), fill = 'grey', aes(alpha = ' Margin (R-S)  '))+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  scale_x_continuous(limits = c(-50, 300), name = 'Safety Margin, MPa')+
  scale_y_continuous(name = 'Probability Density')+
  theme(legend.title = element_blank())+theme(legend.position = 'top')+
  theme(text=element_text(size = 9))


Beta.An <- (mu.R - mu.S) / sqrt(sigma.R^2 + sigma.S^2)
PoF.An <- pnorm(-Beta.An)

```

<div class="notes">
Pr(Fail) = 0.0006871379 (6.87 x 10^-4)
Beta = 3.2
</div>


## Monte Carlo Simulation

```{r, echo=FALSE, warning=FALSE}
N <- 10^3
meanR <- 180; sdR <- 20#; R<-rnorm(N, mean=meanR, sd=sdR)
meanS <- 100; sdS <- 15#; S<-rnorm(N, mean=meanS, sd=sdS)
meanM <- meanR - meanS; sdM <- sqrt(sdR^2 + sdS^2); M <- rnorm(N, mean = meanM, sd = sdM)
#g<-R-S
```



```{r, echo = TRUE, warning=FALSE}
set.seed(123); R <- rnorm(n = N, mean = meanR, sd = sdR)
```

```{r, echo=FALSE, warning=FALSE}
set.seed(1008)
R <- rnorm(N, mean = meanR, sd = sdR); Rcd <- pnorm((R - meanR) / sdR); rDf <- data.frame(R, Rcd)
ggplot(rDf)+ geom_line(aes(y = Rcd, x = R))+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  xlab('Resistance Parameter, R (MPa)') + ylab('Cumulative Distribution Function of R')+
  geom_segment(aes(y = 0.5, yend=0.5, x = 125, xend = meanR), col = 'darkblue',
               arrow = arrow(length = unit(0.25, 'cm'), type = 'closed'), linetype = 5)+
  geom_segment(aes(y = 0.5, yend = 0, x = meanR, xend = meanR), col = 'darkblue',
               arrow = arrow(length = unit(0.25, 'cm'), type='closed'), linetype = 5)

```

## Monte Carlo Simulation

\[
 \small{\Pr(Fail) = \frac  {\sum_{i=1}^{N}{g(u_i)<0}} {N}}
\]

```{r, echo=FALSE, warning=FALSE}

LS<-function(x){
  log( - 0.2 + x/2) - (1.75 + log(sqrt(1.75) / 2))
}

N <- 500; set.seed(1008)
xmc <- rnorm(n = N, mean = 0, sd = 1); ymc <- rnorm(n = N, mean = 0, sd = 1)
xis <- rnorm(n = N, mean = 1.75, sd = 1); yis <- rnorm(n = N, mean = -1.75, sd = 1)

SRAs <- data.frame(xmc, ymc, xis, yis)

ggplot(SRAs)+
  geom_point(aes(x = xmc, y = ymc, col = 'Conventional Monte Carlo'), shape = 1)+
  #geom_point(aes(x=xis, y=yis, col='Importance Sampling'), shape=1)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  scale_x_continuous(limits = c(-5,5), name = 'U1') + scale_y_continuous(limits = c(-5,5), name='U2')+
  #geom_abline(intercept = -3.5, linetype=5)+
  stat_function(fun = LS, linetype = 5, col = 'darkred')+
  #geom_segment(aes(x=0, xend=1.75, y=0, yend=-1.75), col='darkblue',
  #             arrow = arrow(length=unit(0.25, 'cm'), type='closed'))+
  #geom_label(aes(x=3, y=-2.75, label='Reliability \nIndex, Beta'), family='Trebuchet MS', size=3.5)+
  geom_label(aes(x = 0.5, y = -4.75, label = 'g(u)=0'), #family='Trebuchet MS', 
             size = 3.5, col = 'darkred', family = 'Bahnschrift')+
  theme(legend.title = element_blank()) + theme(legend.position = 'top')

```

## Monte Carlo Simulation

<div class="notes">
Calculations using vectors of many realisations of random variables 
</div>

\[
 \small{N_{Min(MC)} = \frac{100}{\Pr(Fail)}}
\]

```{r, echo=FALSE, warning=FALSE, message=FALSE}

table <- read.csv('C:/Users/domen/Desktop/table.csv')
table <- as.data.frame(table)

N.Req <- 100/(mean(table$PoF.An))

#print(table)

MC <- ggplot(table)+
  geom_line(aes(x = table$N, y = table$PoF.An, col='Analytical Solution'))+
  #geom_line(aes(x = table$N, y = table$PoF.lhc, col = 'Latin Hypercube Approximation'))+
  #geom_line(aes(x = table$N, y = table$PoF.is, col = 'Importance Sampling Approximation'))+
  geom_line(aes(x = table$N, y = table$PoF.mc, col = 'Monte Carlo Approximation'))+
  #geom_line(aes(x = table$N, y = table$PoF.lhcis, col = 'Importance Sampling Approximation with LHS'))+
  #scale_linetype_manual(values  =  c(2, 1))+
  geom_vline(xintercept  =  N.Req, col = 'grey', linetype = 5)+
  geom_label(aes(x = N.Req, y = 2*10^-4, label = '145,531 Simulations\n Estimated to be Required'), family = 'Bahnschrift', 
             size = 3.5)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  theme(legend.title  =  element_blank())+theme(legend.position  =  'top')+
  scale_y_continuous(name = 'Probability of Failure')+
  scale_x_continuous(name = 'Log10(Number of Simulations)', trans  =  'log10', breaks = c(10, 10^2, 10^3, 10^4, 10^5))+
  guides(linetype = guide_legend(nrow = 2, byrow  =  TRUE)) + annotation_logticks(sides = 'b')

MC
#library(plotly)
#ggplotly(MC)

```


## Latin Hypercube Sampling

```{r, echo=TRUE, warning=FALSE}
library(lhs)
nVar <- 2; nInt <- 10; A <- randomLHS(nInt,nVar)
```


```{r, echo=FALSE, warning=FALSE, fig.width=4.5, fig.height=4.5, fig.align='center'}
df <- data.frame(A)

xmin <- 0; xmax <- 1; ymin <- 0; ymax <- 1

ggplot(df)+ geom_point(aes(x = A[,1], y = A[,2]), shape = 1, col = 'darkblue')+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  geom_vline(xintercept = 1*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 1*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 2*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 2*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 3*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 3*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 4*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 4*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 5*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 5*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 6*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 6*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 7*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 7*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 8*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 8*(ymax-ymin)/nInt, col='grey', linetype=2)+
  geom_vline(xintercept = 9*(xmax-xmin)/nInt, col='grey', linetype=2)+geom_hline(yintercept = 9*(ymax-ymin)/nInt, col='grey', linetype=2)+
  scale_x_continuous(breaks=seq(0,1,.25), name='Variable 1', limits = c(0, 1))+
  scale_y_continuous(breaks=seq(0,1,.25), name='Variable 2', limits = c(0, 1))

```

## Latin Hypercube Sampling

```{r, echo=FALSE, warning=FALSE, fig.align='center', fig.width=5.5, fig.height=5.5}

nInt <- 10; A <- randomLHS(nInt,nVar); set.seed(1008)
df <- data.frame(nInt = 10, n=paste(nInt, ' samples'), MC = rnorm(n=10), LHS = qnorm(A[,2]))

while (nInt <= 10^4) {
  set.seed(1)
  A <- randomLHS(nInt,nVar)
  MC <- rnorm(n = nInt); LHS <- qnorm(A[,2])
  n <- paste(nInt, 'samples')
  newRow <- data.frame(nInt, n, MC, LHS); df <- rbind(df, newRow); nInt <- nInt * 10; n <- paste(nInt, 'samples')
}

df <- df[-c(1:10), ]

ggplot(df)+
  geom_density(aes(x = MC, alpha = ' Monte Carlo '), fill = 'grey')+
  geom_density(aes(x = LHS, alpha = ' Latin Hypercube '), fill = 'grey')+
  scale_x_continuous(name = 'Random Variable', limits = c(-3,3))+
  scale_y_continuous(name = 'Probability Density')+
  stat_function(geom='line', fun='dnorm', args = list(mean = 0, sd = 1), aes(linetype = ' True Distribution  '))+
  scale_linetype_manual(values = 2)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  theme(legend.title = element_blank())+theme(legend.position = 'top')+
  #guides(alpha=guide_legend(nrow=2, byrow = TRUE))+
  facet_wrap(~n, ncol = 2)

```

## Latin Hypercube Sampling

```{r, echo=FALSE, warning=FALSE}

#setwd('C:/Users/domen/Desktop')
#table<-read.csv('C:/Users/domen/Desktop/table.csv')
table <- as.data.frame(table)

N.Req <- 100 / (mean(table$PoF.An))

#print(table)

ggplot(table)+
  geom_line(aes(x = table$N, y = table$PoF.An, col = 'Analytical Solution'))+
  geom_line(aes(x = table$N, y = table$PoF.lhc, col = 'Latin Hypercube Approximation'))+
  #geom_line(aes(x=table$N, y=table$PoF.is, col='Importance Sampling Approximation'))+
  geom_line(aes(x = table$N, y = table$PoF.mc, col = 'Monte Carlo Approximation'))+
  #geom_line(aes(x=table$N, y=table$PoF.lhcis, col='Importance Sampling Approximation with LHS'))+
  #scale_linetype_manual(values = c(2, 3, 1))+
  #geom_vline(xintercept = N.Req, col='grey', linetype=5)+
  #geom_label(aes(x=N.Req, y=10^-4, label='145,531 Simulations\n Estimated to be Required'), family='Trebuchet MS', size=3.5)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  theme(legend.title = element_blank())+theme(legend.position = 'top')+
  scale_y_continuous(name = 'Probability of Failure')+
  scale_x_continuous(name = 'Number of Simulations', trans = 'log10', breaks = c(10, 10^2, 10^3, 10^4, 10^5))+
  guides(linetype=guide_legend(nrow = 3, byrow = TRUE)) + annotation_logticks(sides = 'b')

```

## Importance Sampling

\[
 \small{\Pr(Fail) = \int_{g(u)<=0} {g(u)\cdot s(u)\cdot w(u) \cdot du}}
\]

```{r, echo=FALSE, warning=FALSE}

LS<-function(x){
  log(-0.2 + x / 2) - (1.75 + log(sqrt(1.75) / 2))
}

N <- 500; set.seed(1008)
xmc <- rnorm(n = N, mean = 0, sd = 1); ymc <- rnorm(n = N, mean = 0, sd = 1)
xis <- rnorm(n = N, mean = 1.75, sd = 1); yis <- rnorm(n = N, mean = -1.75, sd = 1)

SRAs <- data.frame(xmc, ymc, xis, yis)

ggplot(SRAs)+
  geom_point(aes(x = xmc, y = ymc, col = 'Conventional Monte Carlo'), shape = 1)+
  geom_point(aes(x = xis, y = yis, col = 'Importance Sampling'), shape = 1)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  scale_x_continuous(limits = c(-5, 5), name = 'U1') + scale_y_continuous(limits = c(-5,5), name = 'U2')+
  #geom_abline(intercept = -3.5, linetype=5)+
  stat_function(fun = LS, linetype = 5, col = 'darkred')+
  #geom_segment(aes(x=0, xend=1.75, y=0, yend=-1.75), col='darkblue',
  #             arrow = arrow(length=unit(0.25, 'cm'), type='closed'))+
  #geom_label(aes(x=3, y=-2.75, label='Reliability \nIndex, Beta'), family='Trebuchet MS', size=3.5)+
  geom_label(aes(x = 0.5, y = -4.75, label = 'g(u)=0'), #family='Trebuchet MS', 
             size = 3.5, col = 'darkred', family = 'Bahnschrift')+
  theme(legend.title = element_blank()) + theme(legend.position = 'top')

```

## Importance Sampling

```{r, echo=FALSE, warning=FALSE}

#setwd('C:/Users/domen/Desktop')
#table<-read.csv('C:/Users/domen/Desktop/table.csv')
table <- as.data.frame(table)

N.Req <- 100 / (mean(table$PoF.An))

#print(table)

ggplot(table)+
  geom_line(aes(x = table$N, y = table$PoF.An, col = ' Analytical Solution  '))+
  #geom_line(aes(x=table$N, y=table$PoF.lhc, col='Latin Hypercube Approximation'))+
  geom_line(aes(x = table$N, y = table$PoF.is, col = ' Importance Sampling Approximation  '))+
  geom_line(aes(x = table$N, y = table$PoF.mc, col = ' Monte Carlo Approximation  '))+
  #geom_line(aes(x=table$N, y=table$PoF.lhcis, col='Importance Sampling Approximation with LHS'))+
  #scale_linetype_manual(values = c(2, 3, 1))+
  #geom_vline(xintercept = N.Req, col='grey', linetype=5)+
  #geom_label(aes(x=N.Req, y=10^-4, label='145,531 Simulations\n Estimated to be Required'), family='Trebuchet MS', size=3.5)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  theme(legend.title = element_blank()) + theme(legend.position = 'top')+
  scale_y_continuous(name = 'Probability of Failure')+
  scale_x_continuous(name = 'Number of Simulations', trans = 'log10', breaks = c(10, 10^2, 10^3, 10^4, 10^5))+
  guides(linetype = guide_legend(nrow = 3, byrow = TRUE)) + annotation_logticks(sides = 'b')

```

## Importance Sampling

```{r, echo=FALSE, warning=FALSE}

#setwd('C:/Users/domen/Desktop')
#table<-read.csv('C:/Users/domen/Desktop/table.csv')
table <- as.data.frame(table)

N.Req <- 100 / (mean(table$PoF.An))

#print(table)

ggplot(table)+
  geom_line(aes(x = table$N, y=table$PoF.An, col=' Analytical Solution  '))+
  #geom_line(aes(x=table$N, y=table$PoF.lhc, col='Latin Hypercube Approximation'))+
  geom_line(aes(x = table$N, y=table$PoF.is, col = ' Importance Sampling Approximation  '))+
  #geom_line(aes(x=table$N, y=table$PoF.mc, col='Monte Carlo Approximation'))+
  geom_line(aes(x = table$N, y = table$PoF.lhcis, col = ' Importance Sampling with LHS  '))+
  #scale_linetype_manual(values = c(2, 3, 1))+
  #geom_vline(xintercept = N.Req, col='grey', linetype=5)+
  #geom_label(aes(x=N.Req, y=10^-4, label='145,531 Simulations\n Estimated to be Required'), family='Trebuchet MS', size=3.5)+
  theme_ddf_light(base_size = 12, base_family = 'Bahnschrift')+
  theme(legend.title = element_blank()) + theme(legend.position = 'top')+
  scale_y_continuous(name = 'Probability of Failure')+
  scale_x_continuous(name = 'Number of Simulations', trans = 'log10', breaks = c(10, 10^2, 10^3, 10^4, 10^5))+
  guides(linetype = guide_legend(nrow = 3, byrow = TRUE)) + annotation_logticks(sides = 'b')

```

## Thank you for your attention
